High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2021-05-31 Draft_V4046Sgr_mnras.tex
Updates from Overleaf

The observed SED is compared with the model in Fig.~\ref{fig:SED}. From the similarity with the data, we propose that there has to be a small-grain population close to the stars down to 0.3\,au ($R_{\mathrm{sub}}$), i.e. an inner disc, as it was crucial to fit the 5--10\,$\micron$ range. Meanwhile, the decision of employing the three-ringed structure for the small dust (i.e. add Ring5 to the observed structure) relies on the fact that the SED needed a thin ring at a radius at $\sim$5\,au to have a proper fit around 10\,$\micron$, otherwise, instead of the distinct bump that is present, there would be a gap.

@@

Rings, spirals, cavities and gaps are common structures present in protoplanetary discs, but only recent observations have allowed us to study them in detail. Observing systems with solar-type stars, that are not exceptionally bright or large, has not been the main focus of high resolution imaging. For that reason, DARTTS-S programme, that collected DPI data for T Tauri Stars, has been so revealing, and with the counterpart of ALMA continuum maps from DARTTS-A, we can search for substructures in this systems as never before. Following that line, we present new ALMA 1.3\,mm continuum imaging of V4046\,Sgr, a well know circumbinary disc, at an unprecedented definition ($\sim$0$\farcs021 \, \times \, \sim$0$\farcs018$) where new features become visible. Together with the analysis of a SPHERE--IRDIS polarized image and a well-sampled SED, we aim to reproduce the system with radiative transfer modeling looking for a way to explain the data in terms of a physically model.

@@

  \item The central emission present in the ALMA image suggests the existence of a close inner ring that goes in pass the sublimation radius and is made up of dust grains larger than 0.5\,mm. Our interpretation agrees with the mass estimation for this feature made by \citet{Francis_2020} with a mass of 0.012\,M$_{\earth}$.

  \item Our parametric model, which accounts for the SED of the system, involves the existence of sub-micron dust population close to the stars (inner disc outside the sublimation radius). Also predicts the presence of another inner ring at $\sim$5\,au, mainly consisting of small dust grains. The depletion of large dust is consistent with efficient dust trapping in Ring13. This additional ring lies under the coronagraph of the scattered light image and is to faint to be detected by the ALMA observation.

@@

The parametric model presented here has many degrees of freedom and the RT calculation for each set of parameters is very intensive in computational power and time, so a MCMC optimization or similar methods of parameter exploration are impossible to carry out. As a way to compute a measure of goodness of fit and to try to quantify the model uncertainties, we performed a partial exploration of parameter space. We explore $\chi^2$ space in the vicinity of the model values (obtained by trial and error, see Sec.~\ref{sec:model}) for the scale height at $r_\circ$ ($H_\circ$), the width of the gas component of Ring13 $w_{\mathrm{g13}}$, and around the lower limit of the grain sizes in the central blob. The parameter space exploration is shown in Fig.~\ref{fig:chi}. Our model fits the SED and two images, so the total $\chi^2$ value for a given model is composed of the sum of the $\chi^2$ values for each of the three fits.
\begin{equation}
    \chi^2_{\mathrm{total}}=\chi^2_{\mathrm{SED}}+\chi^2_{\mathrm{ALMA}}+\chi^2_{\mathrm{SPHERE}}.
\end{equation}

@@

For the lower grain size limit, we found that despite our chosen value minimises $\chi^2$, the geometry is more suggestive of a border condition. It is evident that there is a threshold around 500\,$\micron$ where the SED start to deviate far away from the typical value. That point represent where the NIR excess becomes significant.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{plot_chi_squared_all800.pdf}
        \caption{One dimensional explorations of the chi squared space of our parametric model. We show the variation of $\chi^2$ in terms of the scale height $H_\circ$ at $r_\circ$=18\,au, the width of the gas component of Ring13 $w_{\mathrm{g13}}$, and around the lower limit of the grain size in the close inner ring. Each panel displays at the bottom the $\chi^2$ values for the SED and for the two images separately, and on top the total $\chi^2$ value in terms of the parameter value.}
    \label{fig:chi}
\end{figure*}

>>>>

The observed SED is compared with the model in Fig.~\ref{fig:SED}. From the similarity with the data, we propose that there has to be a small-grain population close to the stars down to 0.3\,au ($R_{\mathrm{sub}}$), i.e. an inner disc, as it was crucial to fit the 5--10\,$\micron$ range. Meanwhile, the decision of employing the three-ringed structure for the small dust (i.e. add Ring5 to the observed structure) relies on the fact that the SED needs a thin dust ring at a radius of $\sim$5\,au to have a proper fit around 15\,$\micron$, otherwise, instead of the distinct bump that is present, there would be a valley.

@@

Rings, spirals, cavities and gaps are common structures present in protoplanetary discs, but only recent observations have allowed us to study them in detail. In the past years, observing systems with solar-type stars, that are not exceptionally bright or large, has not been the main focus of high resolution imaging. For that reason, DARTTS-S programme, that collected DPI data for T Tauri Stars, has been so revealing, and with the counterpart of ALMA continuum maps from DARTTS-A, we can search for substructures in that kind of systems as never before. Following that line, we present new ALMA 1.3\,mm continuum imaging of V4046\,Sgr, a well know circumbinary disc, at an unprecedented definition ($\sim$0$\farcs021 \, \times \, \sim$0$\farcs018$) where new features become visible. Together with the analysis of a SPHERE--IRDIS polarized image and a well-sampled SED, we aim to reproduce the system with radiative transfer modeling looking for a way to explain the data in terms of a physically model.

@@

  \item The central emission present in the ALMA image suggests the existence of a close inner ring that goes in pass the sublimation radius and is made up of dust grains larger than 0.5\,mm. Our interpretation agrees with the mass estimation of this feature made by \citet{Francis_2020} with a mass of 0.012\,M$_{\earth}$.

  \item Our parametric model, which accounts for the SED of the system, involves the existence of sub-micron dust inner disc. Also predicts the presence of a undetected inner ring at $\sim$5\,au, mainly consisting of small dust grains. The depletion of large dust is consistent with efficient dust trapping in Ring13. This additional ring lies under the coronagraph of the scattered light image and is to faint to be detected by the ALMA observation.

@@

The parametric model presented here has many degrees of freedom and the RT calculation for each set of parameters is very intensive in computational power and time, so a MCMC optimization or similar methods of parameter exploration are impossible to carry out. As a way to compute a measure of goodness of fit and to try to quantify the model uncertainties, we performed a partial exploration of the parameter space.

We explore $\chi^2$ space in the vicinity of the model values (obtained by trial and error, see Sec.~\ref{sec:model}) for the scale height at $r_\circ$ ($H_\circ$), the width of the gas component of Ring13 $w_{\mathrm{g13}}$, and around the lower limit of the grain sizes in the central blob. The parameter space exploration is shown in Fig.~\ref{fig:chi}. Our model fits the SED and two images, so the total $\chi^2$ value for a given model is composed of the sum of the $\chi^2$ values for each of the three fits.
\begin{equation}
    \chi^2_{\mathrm{total}}=\chi^2_{\mathrm{SED}}+\chi^2_{\mathrm{ALMA}}+\chi^2_{\mathrm{SPHERE}}.
\end{equation}

@@

For the lower grain size limit, we found that despite our chosen value seems to minimise $\chi^2$, the geometry is more suggestive of a border condition. So it is evident that there is a threshold around 500\,$\micron$ where the SED starts to deviate strongly from the typical value. That point represent where the NIR excess becomes significant.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{plot_chi_squared_all800.pdf}
        \caption{One dimensional explorations of the chi squared space of our parametric model. The first panel shows the variation of $\chi^2$ in terms of the scale height $H_\circ$ at $r_\circ$=18\,au, the second one for the width of the gas component of Ring13 $w_{\mathrm{g13}}$, and the third one around the lower limit of the grain size in the close inner ring. Each panel displays at the bottom the $\chi^2$ values for the SED and for the two images separately, and on top the total $\chi^2$ value in terms of the parameter value.}
    \label{fig:chi}
\end{figure*}
